---
title: "Ramada by Wyndham Hamilton City Center: Review Benchmarking and Price Prediction"
author: "Sheikh Mohammad Milton, ID No. 1713714"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

set.seed(123) # makes random steps repeatable
ramada_id <- 3572177  # Assigning Ramada's Tripadvisor id

```

#1. Introduction

## 1.1 Purpose of the analysis

The purpose of this assignment is to conduct a comprehensive, data-driven analysis and develop a machine learning model to support strategic decision-making for the manager of Ramada by Wyndham Hamilton City Center. 

The manager of **Ramada by Wyndham Hamilton City Center** (Tripadvisor id 3572177) wants to know:

1. What are the hotel's **strengths and weaknesses**, and how does it compare with **competitors**?

2. What price per night would a **machine learning model** estimate for the hotel?

3. What should the hotel **do to improve**, and should the current average price of **$183.50** change?

For finding out the answers of the above question following analysis has been presented.

## 1.2 Load packages

```{r initial}
# Install and load packages

if(!require("pacman"))install.packages("pacman")

pacman::p_load(readr, dplyr, tidyr, stringr, lubridate, purrr, ggplot2, gridExtra, GGally, corrplot)
pacman::p_load(tidytext, textstem, wordcloud, textdata, topicmodels)
pacman::p_load(DescTools, skimr, finalfit, rstatix, huxtable, jtools)
pacman::p_load(knitr, easystats, gtsummary, gt, scales)
pacman::p_load(SnowballC, textstem)
```

# 2. Data Import and Initial Inspection

## 2.1 Import the dataset

```{r}
hotel_main <- read_csv("week14_hotelmain.csv")
hotel_pred <- read_csv("week14_hotelpred.csv")
hotel_review <- read_csv("week14_review.csv")
```

## 2.2 Initial Inspection

```{r}
# How many rows and columns in each file?
tibble(file = c("hotel_main", "hotel_pred", "hotel_review"),
       rows = c(nrow(hotel_main), nrow(hotel_pred),
                nrow(hotel_review)),
       columns = c(ncol(hotel_main), ncol(hotel_pred),
                   ncol(hotel_review)))

# Variable names, types and first values
glimpse(hotel_main)
glimpse(hotel_review)

# A look at a few rows (only key columns, so the table is readable)
hotel_main |>
  select(id, name, category, locality, hotel_class, price_class, price, room, review, rating_overall) |>
    head(5)

hotel_pred |>
  select(id, name, category, locality, hotel_class, price_class, room, review, rating_overall)

hotel_review |>
  select(hotel_id, review_date, trip_type, rating_overall, title) |>
  head(5)

# A quick look of duplicate rows

cat("Duplicate rows in hotel_main:", sum(duplicated(hotel_main)), "\n")

cat("Duplicate rows in hotel_review:", sum(duplicated(hotel_review)), "\n")

cat("Ramada in hotel_main? ", ramada_id %in% hotel_main$id, "\n")

cat("Ramada in hotel_pred? ", ramada_id %in% hotel_pred$id, "\n")

cat("Ramada reviews in hotel_review:", sum(hotel_review$hotel_id == ramada_id), "\n")

cat("Review date range:", format(min(hotel_review$review_date)), "to", format(max(hotel_review$review_date)), "\n")

# Basic summary of the main numeric variables
hotel_main |>
  select(price, room, photo, review, hotel_class, rating_overall) |>
  summary()
```

## 2.3 Initial Findings

* **Size:** `hotel_main` has 729 rows and 47 columns (one row per accommodation), `hotel_pred` has a single row (Ramada) and 46 columns, and `hotel_review` has 9,585 reviews and 19 columns.

* **Ramada is only in `hotel_pred`**, not in `hotel_main`, so the price model never sees Ramada while learning. Its 349 reviews are in `hotel_review` (Tripadvisor shows 360 reviews in total on the hotel page).

* **Duplicates:** 8 duplicated hotel rows and 1 duplicated review. They will be removed before further analysis.

* **Price is very skewed:** the median is about $190 but the mean is about $242 and the maximum is $4,014.50, so a few luxury lodges pull the average up.

* **`hotel_class` has many zeros** (more than half of the hotels). A hotel cannot have zero stars, so 0 means "not rated", which we handle in Phase 2.

* **Amenity columns contain only 1 or `NA`**, so `NA` means "not listed".

* Reviews run from February 2005 to July 2026, so we can study trends over time.

* Some categorical data has been treated as CHARACTER Data Type which have to be converted to CATEGORICAL Data before analysis.

# 3. Data perpetration

## 3.1 Missing value management

Missing values (`NA`) can break models and bias averages. First we measure how much is missing in each variable, then we decide purposefully what to do for each group of variables (not just delete everything).

```{r missing-main}
# Percentage of missing values in each column of hotel_main

missing_main <- hotel_main |>
  summarise(across(everything(), ~ mean(is.na(.x)) * 100)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "pct_missing") |>
  filter(pct_missing > 0) |>
  arrange(desc(pct_missing))

print(missing_main, n = Inf)
```

```{r missing-main}
# Number and percentage of missing values in each column of hotel_main

missing_main <- tibble(
  variable = names(hotel_main),
  n_missing = colSums(is.na(hotel_main)),
  pct_missing = round(colMeans(is.na(hotel_main)) * 100, 1)) |>
  filter(n_missing > 0) |>
  arrange(desc(n_missing))

kable(missing_main,
      col.names = c("Variable", "Missing values (count)", "Missing (%)"),
      caption = paste("Missing values in hotel_main (total rows:", nrow(hotel_main), ")"))
```

**Interpretation:**

* The amenity columns have the most missing values (about 48% to 94%). Because they only contain 1 or `NA`, a missing value really means "the amenity is not listed", so replacing it with 0 is sensible.

* Price and price class are each missing for 39% of hotels, and they are missing for exactly the same 286 hotels. Those hotels cannot be used to train a price model. It also means `price_class` is built from price, so we must not use it to predict price (leakage).

* Offer prices are about 68% missing and are almost identical to price, so they are not used as predictors.

* Ratings are missing for 16% to 28% of hotels, mostly hotels with no reviews at all.

* In the review file, `rating_overall` is never missing but the sub-ratings (cleanliness, room, sleep, value, location) are missing for 54% to 59% of reviews, because guests often leave only an overall rating. `hotel_response` is missing for 49% of reviews, which simply means the hotel did not reply.

```{r missing-review}
# Missing values in the review file
hotel_review |>
  summarise(across(everything(), ~ mean(is.na(.x)) * 100)) |>
  pivot_longer(everything(), names_to = "variable",
               values_to = "pct_missing") |>
  filter(pct_missing > 0) |>
  arrange(desc(pct_missing))
```

## 3.2 Removing duplicates

```{r clean-data}
# 1. Remove duplicates
hotel_main <- hotel_main |> distinct()
hotel_review <- hotel_review |> distinct()

# The amenity columns
amenity_cols <- names(hotel_main)[str_starts(names(hotel_main), "amenity_")]

# The most common localities (the others are grouped as "Other")
top_localities <- hotel_main |>
  filter(!is.na(locality)) |>
  count(locality, sort = TRUE) |>
  slice_head(n = 10) |>
  pull(locality)

top_localities
```

## 3.3 Creating new variables

New variables simplify the information (for example "Economy / Moderate / Premium / Luxury" is easier to compare than "$", "$$" and so on), and some capture ideas such as how many amenities a hotel offers. We write one small function so that exactly the same steps are applied to `hotel_main` and to Ramada's row in `hotel_pred`.

```{r new-variables-hotel}
add_hotel_vars <- function(df) {
  df |>
    mutate(

      # missing locality -> "Unknown"
      locality = replace_na(locality, "Unknown"),
      
      # amenity NA -> 0 (not listed), then count how many amenities
      across(all_of(amenity_cols), ~ replace_na(as.numeric(.x), 0)), amenity_count = rowSums(across(all_of(amenity_cols))),
      
      # price category from price_class
      price_cat = case_when(
        price_class == "$"    ~ "Economy",
        price_class == "$$"   ~ "Moderate",
        price_class == "$$$"  ~ "Premium",
        price_class == "$$$$" ~ "Luxury",
        TRUE                  ~ "Unknown"),
      price_cat = factor(price_cat,
                         levels = c("Economy", "Moderate", "Premium", "Luxury", "Unknown")),

      # star rating: 0 means unrated
      star_num = na_if(hotel_class, 0),
      star_cat = case_when(
        is.na(star_num) ~ "Unrated",
        star_num < 3    ~ "1-2.5 Stars",
        star_num < 4    ~ "3-3.5 Stars",
        TRUE            ~ "4-5 Stars"),
      star_cat = factor(star_cat,
                        levels = c("1-2.5 Stars", "3-3.5 Stars", "4-5 Stars", "Unrated")),

      # size of the hotel
      room_size = case_when(
        is.na(room) ~ "Unknown",
        room <= 10  ~ "Small (1-10)",
        room <= 30  ~ "Medium (11-30)",
        TRUE        ~ "Large (31+)"),
      room_size = factor(room_size,
                         levels = c("Small (1-10)", "Medium (11-30)", "Large (31+)", "Unknown")),

      # position in the destination ranking (small = better), as a proportion
      rank_pct = rank_position / rank_denominator,

      # grouped location
      area = if_else(locality %in% top_localities, locality, "Other")
    )
}

hotel_main <- add_hotel_vars(hotel_main)
hotel_pred <- add_hotel_vars(hotel_pred)

```


```{r}
# One table with all hotels (Ramada included) for benchmarking
hotel_all <- bind_rows(hotel_main, hotel_pred)

hotel_main |>
  select(name, price_class, price_cat, hotel_class, star_cat, room, room_size, amenity_count, area) |>
  head(10)
```


```{r new-variables-review}
# Variables for the review file
hotel_review <- hotel_review |>
  left_join(hotel_all |>
              select(id, hotel_name = name, locality, category),
            by = c("hotel_id" = "id")) |>
  mutate(
    review_year = year(review_date),
    responded = if_else(is.na(hotel_response), 0, 1),            # did the hotel reply?
    word_count  = str_count(text, "\\S+"),                       # length of the review
    trip_type = replace_na(str_to_title(trip_type), "Not stated"),
    trip_type = if_else(trip_type == "None", "Not stated", trip_type),
    origin = case_when(
      is.na(user_location)  ~ "Unknown",
      str_detect(user_location, "New Zealand") ~ "New Zealand",
      str_detect(user_location, "Australia") ~ "Australia",
      TRUE  ~ "Other countries"),
    full_text = paste(title, text),                             # title + text for text mining
    sentiment_group = case_when(
      rating_overall >= 4 ~ "Positive (4-5)",
      rating_overall == 3 ~ "Neutral (3)",
      TRUE                ~ "Negative (1-2)"),
    # who is who?
    group = case_when(
      hotel_id == ramada_id ~ "Ramada",
      locality == "Hamilton" ~ "Hamilton competitors",
      TRUE ~ "Other areas")
  )

# How many reviews per group?
hotel_review |> count(group)
```

## 3.4 Defining Competitors

Ramada's competitors are the other hotels located in Hamilton. We use the same competitors for both the hotel-level and review-level comparisons.

```{r define-groups}
rating_vars <- c("rating_overall", "rating_cleanliness", "rating_location", "rating_room", "rating_service", "rating_sleep", "rating_value")

ramada_row <- hotel_all |> filter(id == ramada_id)
ham_comp   <- hotel_all |> filter(locality == "Hamilton", id != ramada_id)

comparison_hotel <- tibble(
  aspect       = rating_vars,
  Ramada       = unlist(ramada_row[rating_vars]),
  Hamilton_avg = sapply(ham_comp[rating_vars], mean, na.rm = TRUE)
)

ramada_reviews <- hotel_review |> filter(group == "Ramada")
comp_reviews <- hotel_review |> filter(group == "Hamilton competitors")

cat("Hamilton competitor hotels:", nrow(ham_comp), "\n")
cat("Competitor hotels with reviews:", n_distinct(comp_reviews$hotel_id), "\n")
```

# 4. Exploratory Data Analysis (EDA)

## 4.1 Understanding Price

Price is what we will predict through Machine Learning, so we must understand its typical level, how spread out it is, and whether there are extreme values.

```{r price-statistics}
# Overall price
hotel_main |>
  filter(!is.na(price)) |>
  summarise(n = n(), mean = mean(price), median = median(price), sd = sd(price), min = min(price), max = max(price))

# By type of accommodation
hotel_main$category <- as.factor(hotel_main$category)
hotel_main |>
  filter(!is.na(price)) |>
  group_by(category) |>
  summarise(n = n(), mean = round(mean(price)), median = median(price), sd = round(sd(price)), min = min(price), max = max(price))

# By price category
hotel_main |>
  filter(!is.na(price)) |>
  group_by(price_cat) |>
  summarise(n = n(), mean = round(mean(price)), median = median(price))

# By star category
hotel_main |>
  filter(!is.na(price)) |>
  group_by(star_cat) |>
  summarise(n = n(), mean = round(mean(price)), median = median(price))
```

```{r price-outliers}
# Outliers using the IQR rule: values above Q3 + 1.5 x IQR
price_vals <- hotel_main$price[!is.na(hotel_main$price)]
q1 <- quantile(price_vals, 0.25)
q3 <- quantile(price_vals, 0.75)
upper_fence <- q3 + 1.5 * (q3 - q1)

cat("Upper fence:", round(upper_fence), "\n")
cat("Hotels above the fence:", sum(price_vals > upper_fence), "\n")
cat("Hotels above $1000:", sum(price_vals > 1000), "\n")

hotel_main |>
  filter(!is.na(price)) |>
  arrange(desc(price)) |>
  select(name, locality, category, price, room) |>
  head(10)
```

## 4.2 Ramada's current price versus Hamilton hotels

```{r price-hamilton}
ham_price <- ham_comp |> filter(!is.na(price))

ham_price |>
  summarise(hotels = n(), 
            mean = round(mean(price)), 
            median = median(price), 
            q25 = quantile(price, 0.25), 
            q75 = quantile(price, 0.75))

cat("Share of Hamilton competitors priced at or below $183.50:", round(mean(ham_price$price <= 183.5) * 100), "%\n")

ham_price |>
  group_by(category) |>
  summarise(hotels = n(), median = median(price))
```

Findings:

* Price is strongly right-skewed: the mean (about $243) is well above the median (about $189), and the standard deviation (about $283) is larger than the mean. Extreme lodges (for example $4,014 and $2,528) pull the average up, so the median is a better "typical" price.

* By type: Hotels have the narrowest spread (standard deviation about $94). Specialty Lodging has the highest mean ($309) but a much lower median ($206), which shows a few very expensive properties. Bed and Breakfasts are in between.

* By price category: the median rises clearly from Economy ($139) to Moderate ($220), Premium ($378) and Luxury ($884).

* By stars: 4-5 star hotels have a median of about $230 compared with about $175 for 3-3.5 star hotels, so star class matters for price.

* Outliers: 33 hotels are above the IQR upper fence of about $403 and 9 are above $1,000. These are luxury lodges, very different from a city hotel such as Ramada, so we exclude hotels above $1,000 when training the model.

* Ramada's position: its current price of $183.50 is close to the overall median, but in Hamilton it is above most competitors. The median Hamilton price is about $148 and Ramada is priced higher than roughly 63% of Hamilton hotels. Only the Hamilton hotels classified as "Moderate" by Tripadvisor have a higher median (about $208), so Ramada is already priced in the upper-middle of its local market.

## 4.3 Graphs and interpretation

### Graph 1: How is price distributed?

A histogram shows the shape of the price data. Right-skewed data (a long tail of expensive hotels) is usually easier to model after a log transformation.

```{r g-price-hist}
g1 <- hotel_main |>
  filter(!is.na(price)) |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 40, fill = "steelblue", colour = "white") +
  geom_vline(xintercept = 183.5, colour = "red", linetype = "dashed") +
  labs(title = "Distribution of price per night (red line = Ramada's $183.50)",
       x = "Price per night (NZD)", y = "Number of hotels")

g2 <- hotel_main |>
  filter(!is.na(price)) |>
  ggplot(aes(x = log(price))) +
  geom_histogram(bins = 40, fill = "darkorange", colour = "white") +
  labs(title = "Distribution of log(price)", x = "log(price)", y = "Number of hotels")

grid.arrange(g1, g2, ncol = 2)
```

**Interpretation:**

* The first histogram shows a long right tail: most hotels charge less than about $400, but a handful charge far more. Skewness (a measure of lopsidedness) is about 8.0 for price.

* The red dashed line shows Ramada's $183.50, which is near the middle of the distribution.

* After taking the log, the shape is much more balanced (skewness falls to about 1.0). 

### Graph 2: Price by type of accommodation and by price category

Boxplots compare the middle and spread of price across groups. We use a log axis because of the long tail.

```{r g-price-box}
hotel_main |>
  filter(!is.na(price)) |>
  ggplot(aes(x = category, y = price, fill = category)) +
  geom_boxplot(show.legend = FALSE) +
  scale_y_log10(labels = dollar) +
  labs(title = "Price by type of accommodation (log scale)", x = NULL, y = "Price per night")

hotel_main |>
  filter(!is.na(price)) |>
  ggplot(aes(x = star_cat, y = price, fill = star_cat)) +
  geom_boxplot(show.legend = FALSE) +
  scale_y_log10(labels = dollar) +
  labs(title = "Price by star category (log scale)", x = NULL, y = "Price per night")
```

**Interpretation:**

* By type: the medians for Hotels, Bed and Breakfasts and Specialty Lodging are fairly close (about $178 to $206). Specialty Lodging has the widest spread and the most extreme points, so type alone does not set the price level.

* By stars: price rises with star class. 4-5 star properties have the highest median, 1-2.5 stars the lowest (about $113). The "Unrated" group is a mixture of all kinds of properties and has a wide spread.

* Conclusion: star class is a more useful price signal than accommodation type.

### Graph 3: Does a better rating mean a higher price?

A scatter plot shows the relationship between two numeric variables. If guests rate a hotel higher, can it charge more?

```{r g-price-rating}
hotel_main |>
  filter(!is.na(price), !is.na(rating_overall)) |>
  ggplot(aes(x = rating_overall, y = price, colour = category)) +
  geom_jitter(alpha = 0.6, width = 0.03) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, colour = "black") +
  scale_y_log10(labels = dollar) +
  labs(title = "Price versus overall rating (log price axis)",
       x = "Overall rating (out of 5)", y = "Price per night", colour = "Type")
```

**Interpretation:**

* The black line slopes upwards: hotels with higher guest ratings tend to charge more.

* However, the points are widely scattered. Hotels with the same rating can have very different prices, so rating explains only part of the price. Other factors such as location, size, season and facilities also matter.

* Ramada's overall rating (3.0) is at the low end of the range. If guests rate a hotel lower, the market supports a lower price, which is relevant to the manager's plan to raise the price.

### Graph 4: Correlation heatmap

Correlation measures how strongly two numeric variables move together (from -1 to +1). The heatmap shows which variables are related to price and which are related to each other.

```{r g-corr, fig.height=7, fig.width=7}
cor_data <- hotel_main |>
  select(price, room, photo, review, star_num, rank_pct, amenity_count,
         all_of(rating_vars))

cor_matrix <- cor(cor_data, use = "pairwise.complete.obs")

corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8, addCoef.col = "black", number.cex = 0.6, tl.col = "black")
```

**Interpretation:**

* Relationships with price are modest. The strongest correlations with price are star class (about 0.35), the room rating (0.23), the overall rating (0.22) and the sleep rating (0.21). Number of reviews (0.00), number of amenities (0.01) and number of rooms (-0.06) show almost no linear relationship with price. (Extreme prices weaken these correlations, which is another reason to treat the biggest outliers carefully.)

* The rating variables are strongly correlated with one another (about 0.78 to 0.86, with location lower at 0.58 to 0.71), and the number of photos and number of reviews are almost the same signal (0.88). They carry overlapping information, so a simple linear model may struggle with them (multi-collinearity). Regularized models (lasso/ridge) and tree-based models cope better

### Graph 5: Where are the expensive hotels?

Location is a classic price driver. We look at average price by area and at a simple map of hotels coloured by price.

```{r g-area}
hotel_main |>
  filter(!is.na(price)) |>
  group_by(area) |>
  summarise(median_price = median(price), n = n()) |>
  ggplot(aes(x = reorder(area, median_price), y = median_price)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0("n=", n)), hjust = -0.1, size = 3) +
  coord_flip() +
  scale_y_continuous(labels = dollar, expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Median price per night by area", x = NULL, y = "Median price")

hotel_main |>
  filter(!is.na(price), price <= 1000) |>
  ggplot(aes(x = longitude, y = latitude, colour = price)) +
  geom_point(alpha = 0.8) +
  scale_colour_viridis_c(trans = "log10", labels = dollar) +
  labs(title = "Hotel locations coloured by price", colour = "Price")

# library(ggplot2)
# library(dplyr)
# install.packages("sf")
# library(sf)
# install.packages("rnaturalearth")
# library(rnaturalearth)
# library(scales) # Required for the 'dollar' label function
# 
# # 1. Fetch a background world map
# world_map <- ne_countries(scale = "medium", returnclass = "sf")
# 
# # 2. Clean and convert your hotel data to a spatial (sf) object
# hotel_map_data <- hotel_main |>
#   filter(!is.na(price), price <= 1000) |>
#   # Crucial step: Converts lat/long columns into an sf geometry column
#   st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)
# 
# # 3. Build the map with ggplot2
# ggplot() +
#   # Draw the background map layer first
#   geom_sf(data = world_map, fill = "gray95", color = "lightgray", size = 0.2) +
#   
#   # Layer your hotel points on top using geom_sf
#   geom_sf(data = hotel_map_data, aes(colour = price), alpha = 0.8, size = 1.5) +
#   
#   # Format the colour gradient and scales
#   scale_colour_viridis_c(trans = "log10", labels = label_dollar()) +
#   
#   # Automatically frame the map boundaries to your specific data points
#   coord_sf(xlim = range(hotel_main$longitude, na.rm = TRUE), 
#            ylim = range(hotel_main$latitude, na.rm = TRUE)) +
#   
#   # Labels and themes
#   labs(
#     title = "Hotel Locations Coloured by Price",
#     x = "Longitude", 
#     y = "Latitude",
#     colour = "Price"
#   ) +
#   theme_minimal()

```

**Interpretation:**

* Hamilton, Ramada's own market, has the lowest median price (about $148, 67 hotels), because it is a large city with many motels and business hotels. Tourist destinations are more expensive: Raglan (about $297, but only 10 hotels), Whitianga (about $213), Taupo (about $203) and Cambridge (about $200).

* The map shows that hotels cluster around towns, and the colour pattern confirms that location is related to price. This is why `area`, latitude and longitude are included in the model.

* Ramada's $183.50 is therefore a relatively high price for Hamilton.

### Graph 6: Ramada versus Hamilton competitors on each aspect

This is the key benchmarking graph. Bars side by side show where Ramada is above or below the Hamilton average.

```{r g-ramada-vs-comp}
comparison_hotel |>
  pivot_longer(c(Ramada, Hamilton_avg), names_to = "who", values_to = "score") |>
  mutate(aspect = str_remove(aspect, "rating_"),
         who = recode(who, Hamilton_avg = "Hamilton average")) |>
  ggplot(aes(x = aspect, y = score, fill = who)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = round(score, 1)), position = position_dodge(width = 0.9),
            vjust = -0.3, size = 3) +
  scale_fill_manual(values = c("Ramada" = "firebrick", "Hamilton average" = "grey60")) +
  coord_cartesian(ylim = c(0, 5.2)) +
  labs(title = "Ramada vs Hamilton competitors (Tripadvisor hotel-level scores)",
       x = NULL, y = "Average rating (out of 5)", fill = NULL)
```

**Interpretation:**

* Ramada (red) is **lower than the Hamilton average (grey) on every aspect except location**.
* The biggest gaps are overall rating, room, value and sleep. These are the aspects that guests link most closely to the quality of the physical product and the price paid.
* Location (4.4) is Ramada's strongest score and the main competitive advantage.

### Graph 7: Distribution of star ratings in reviews

**Why are we doing this?**
An average can hide the detail. Looking at the *share* of 1 to 5 star reviews shows whether guests are polarised or consistently happy.

```{r g-rating-dist}
hotel_review |>
  filter(group %in% c("Ramada", "Hamilton competitors")) |>
  count(group, rating_overall) |>
  group_by(group) |>
  mutate(pct = n / sum(n)) |>
  ggplot(aes(x = factor(rating_overall), y = pct, fill = group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("Ramada" = "firebrick", "Hamilton competitors" = "grey60")) +
  labs(title = "Share of reviews by star rating",
       x = "Overall rating given by the guest", y = "% of reviews", fill = NULL)
```

**Interpretation:**

* Ramada's review ratings are polarised and skewed to the low end: about 37% of its reviews are 1 or 2 stars (21% give 1 star), compared with about 13% for competitors. Only about 39% of Ramada reviews are 4 or 5 stars, compared with about 73% for competitors.

* Five-star reviews make up 19% of Ramada's reviews but 41% of competitors' reviews.

* This means the low average is not caused by a few angry guests. A large share of guests have a poor experience.

### Graph 8: Has Ramada's rating changed over time?

A trend line shows whether the hotel is improving or getting worse. Years with fewer than 5 reviews are dropped because the average would be unreliable.

```{r g-trend}
hotel_review |>
  filter(group %in% c("Ramada", "Hamilton competitors"), review_year >= 2012) |>
  group_by(group, review_year) |>
  summarise(avg_rating = mean(rating_overall), n = n(), .groups = "drop") |>
  filter(n >= 5) |>
  ggplot(aes(x = review_year, y = avg_rating, colour = group)) +
  geom_line(linewidth = 1) +
  geom_point(aes(size = n)) +
  scale_colour_manual(values = c("Ramada" = "firebrick", "Hamilton competitors" = "grey50")) +
  labs(title = "Average review rating by year (point size = number of reviews)",
       x = "Year of review", y = "Average overall rating", colour = NULL, size = "Reviews")
```

**Interpretation:**

* Ramada is below the competitor average in every year with enough reviews (2013 to 2025). Competitors stay around 3.6 to 4.0, while Ramada ranges from about 2.3 to 3.7.

* There is no clear sustained improvement: for example Ramada averaged about 3.3 in 2019, 3.2 in 2022, 2.7 in 2023 and 3.2 in 2024. The gap narrowed in some years but never closed.

* Be careful with 2025 and 2026: there are only 10 and 8 Ramada reviews, so those averages are unreliable. There are too few Ramada reviews in 2020 to plot, which fits the COVID period.

### Graph 9: Ramada ratings by type of traveler

Different guests want different things (business travellers vs families). Knowing which segment is least satisfied helps target improvements.

```{r g-trip}
ramada_reviews |>
  group_by(trip_type) |>
  summarise(avg_rating = mean(rating_overall), n = n()) |>
  ggplot(aes(x = reorder(trip_type, avg_rating), y = avg_rating)) +
  geom_col(fill = "firebrick") +
  geom_text(aes(label = paste0(round(avg_rating, 2), " (n=", n, ")")), hjust = -0.05, size = 3.5) +
  coord_flip(ylim = c(0, 5)) +
  labs(title = "Ramada: average rating by trip type", x = NULL, y = "Average overall rating")
```

**Interpretation:**

* Business travellers (average about 2.69, 85 reviews), friends (2.58, 26 reviews) and solo travellers (2.77, 13 reviews) give the lowest ratings. Couples (3.21) and families (3.12) are somewhat happier.

* Business guests are the largest identifiable group of reviewers and one of the least satisfied, which matters because they are probably a core market for a city-centre hotel.

* The friends and solo groups are small, so treat those two bars with caution.

### Graph 10: Does the hotel respond to reviews?

Replying to guests (especially unhappy ones) shows that the hotel cares and can protect its reputation. We compare response rates by rating.

```{r g-response}
hotel_review |>
  filter(group %in% c("Ramada", "Hamilton competitors")) |>
  group_by(group, sentiment_group) |>
  summarise(response_rate = mean(responded), n = n(), .groups = "drop") |>
  ggplot(aes(x = sentiment_group, y = response_rate, fill = group)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("Ramada" = "firebrick", "Hamilton competitors" = "grey60")) +
  labs(title = "Share of reviews that received a hotel response",
       x = "Review type", y = "Response rate", fill = NULL)
```

**Interpretation:**

* Ramada replies to 85% to 87% of reviews whatever the rating. Competitors reply to only about 32% of negative reviews, 49% of neutral reviews and 58% of positive reviews.

* Ramada's responsiveness is a real strength: it shows guests that the management is listening. However, replying does not by itself fix the underlying problems that cause the low ratings.

## 4.4 Statistical tests

A graph can suggest a difference, but a statistical test tells us whether the difference is real or could just be chance. We use a significance level of 5%: if the p-value is below 0.05 we say the result is statistically significant. Because price and ratings are skewed or ordinal, we mainly use non-parametric tests (Kruskal-Wallis, Wilcoxon, Spearman) that do not assume a normal distribution.

### Test 1: Does price differ across types of accommodation? (Kruskal-Wallis)

* H0: median price is the same for Hotels, Bed and Breakfasts (B&Bs) and Specialty Lodging.
* H1: at least one type has a different price.

```{r test-kruskal-price}
price_data <- hotel_main |> filter(!is.na(price))

kruskal.test(price ~ category, data = price_data)

# Which pairs differ? (Bonferroni-adjusted p-values)
pairwise.wilcox.test(price_data$price, price_data$category, p.adjust.method = "bonferroni")
```

**Interpretation:**

* The Kruskal-Wallis p-value is 0.039, which is below 0.05, so we reject H0: price is not the same across all types of accommodation.

* The pairwise tests show that the only significant difference is Hotel vs Specialty Lodging (adjusted p = 0.036). Bed and Breakfasts are not significantly different from the others.

* The evidence is moderate: the type of accommodation matters a little, but much less than star class.

### Test 2: Which variables are related to price? (Spearman correlation)

* H0: there is no monotonic relationship between price and the variable (correlation = 0).

```{r test-spearman}
test_vars <- c("rating_overall", "rating_value", "rating_room", "room", "photo", "review", "amenity_count", "star_num")

spearman_table <- map_dfr(test_vars, function(v) {
  ok <- complete.cases(price_data$price, price_data[[v]])
  t <- cor.test(price_data$price[ok], price_data[[v]][ok], method = "spearman", exact = FALSE)
  tibble(variable = v, n = sum(ok), rho = round(unname(t$estimate), 3), p_value = signif(t$p.value, 3))
})

spearman_table
```

**Interpretation:**

* The strongest positive relationships with price are star class (rho about 0.50), the room rating (about 0.50), the overall rating (about 0.43) and the value rating (about 0.33). All have p-values far below 0.05.

* The number of rooms has a significant negative relationship (rho about -0.20): larger properties tend to charge less per night.

* Photos (0.16) and number of amenities (0.10) are statistically significant but very weak. The number of reviews is not significant (p = 0.30), which means popularity does not translate into price.

* Quality indicators (stars and room quality) are what the market pays for. This is important for Ramada, whose room rating (3.2) is low.

### Test 3: Is the type of accommodation linked to its price level? (Chi-square)

* H0: type of accommodation and price level are independent.

```{r test-chisq}
chi_data <- hotel_main |>
  filter(price_cat != "Unknown") |>
  mutate(price_level = if_else(price_cat %in% c("Premium", "Luxury"), "Premium/Luxury", as.character(price_cat)))

chi_table <- table(chi_data$category, chi_data$price_level)
chi_table
chisq.test(chi_table)
```

**Interpretation:**

* The chi-square p-value is 0.002 (below 0.05), so we reject independence: the type of accommodation is linked to its price level.

* Specialty Lodging is much more likely to be Premium/Luxury (26 of 134 properties, about 19%) than Hotels (6 of 151, about 4%). Hotels are mostly Economy or Moderate.

* This fits the earlier boxplots: very expensive properties are mostly specialty lodges, not hotels like Ramada.

### Test 4: Is Ramada's rating different from competitors? (Wilcoxon rank-sum)

* H0: Ramada and Hamilton competitors have the same rating distribution.

* We run the test for 7 ratings, so we use the Bonferroni rule: significant only if p < 0.05 / 7 = 0.007. This avoids false alarms from running many tests.

```{r test-wilcoxon}
both <- hotel_review |> filter(group %in% c("Ramada", "Hamilton competitors"))

wilcoxon_table <- map_dfr(rating_vars, function(v) {
  ok <- both |> filter(!is.na(.data[[v]]))
  t <- wilcox.test(ok[[v]] ~ ok$group, exact = FALSE)
  tibble(
    rating      = v,
    Ramada_mean = round(mean(ok[[v]][ok$group == "Ramada"]), 2),
    Comp_mean   = round(mean(ok[[v]][ok$group == "Hamilton competitors"]), 2),
    p_value     = signif(t$p.value, 3)
  )
}) |>
  mutate(difference = Ramada_mean - Comp_mean,
         verdict = case_when(
           p_value >= 0.05 / 7 ~ "No significant difference",
           difference < 0      ~ "Ramada significantly LOWER",
           TRUE                ~ "Ramada significantly HIGHER"))

wilcoxon_table
```

**Interpretation:**

* Using the stricter Bonferroni threshold (0.007), Ramada is significantly lower than competitors on six of the seven ratings, with p-values far below 0.001.

* The gaps are largest for the overall rating (-0.92), room (-0.81) and value (-0.81), then sleep (-0.68), cleanliness (-0.59) and service (-0.58).

Location is the only rating with no significant difference (4.39 for both, p = 0.79).

* So Ramada's weaknesses are real and not due to chance, and location is its one area of parity with competitors.

### Test 5: Do traveller types rate Ramada differently? (Kruskal-Wallis)

```{r test-trip}
kruskal.test(rating_overall ~ trip_type, data = ramada_reviews)
```

### Test 6: Has Ramada improved since the COVID period? (Wilcoxon)

* Compare reviews posted up to 2019 with reviews posted from 2022 onward.

```{r test-improve}
ramada_period <- ramada_reviews |>
  filter(review_year <= 2019 | review_year >= 2022) |>
  mutate(period = if_else(review_year <= 2019, "Up to 2019", "2022 onward"))

ramada_period |> group_by(period) |> summarise(n = n(), mean_rating = round(mean(rating_overall), 2))

wilcox.test(rating_overall ~ period, data = ramada_period, exact = FALSE)
```

**Interpretation of Tests 5 and 6:**

* Test 5: p = 0.052, just above 0.05, so the differences between traveller types are suggestive but not statistically significant. Business travellers look less satisfied, but we cannot be certain.

* Test 6: the average rating was 2.99 up to 2019 and 3.04 from 2022 onward, with p = 0.68. There is no significant improvement since the COVID period, so the problems have not been solved over time.

### Tests 7 to 10: Parametric tests (t-test, ANOVA, Pearson correlation)

So far we used non-parametric tests, which compare ranks. Parametric tests compare means and are more powerful when their assumptions are reasonable. Their main assumption is roughly normal data, or a large sample size (the Central Limit Theorem makes means behave normally when samples are large). Here we use them where this holds:

* review ratings, where we have hundreds of reviews per group, and
* log(price), which is much more symmetric than price (see Graph 1).

When we compare means of `log(price)`, a difference can be turned into a percentage: `exp(difference) - 1`.

#### Test 7: Do Ramada and competitors have different average review ratings? (Welch t-test)

* H0: the mean overall rating is the same for Ramada and Hamilton competitors.
* H1: the means are different. 

```{r test-ttest-rating}
t.test(rating_overall ~ group, data = both)
```

#### Test 8: Are Hamilton hotels cheaper than hotels in other areas? (Welch t-test on log price)

* H0: mean log(price) is the same in Hamilton and elsewhere.

```{r test-ttest-price}
price_area <- price_data |>
  mutate(market = if_else(locality == "Hamilton", "Hamilton", "Other areas"))

t.test(log(price) ~ market, data = price_area)

# "Typical price" = exp(mean of log price), which is the geometric mean
price_area |>
  group_by(market) |>
  summarise(hotels = n(), typical_price = round(exp(mean(log(price)))))
```

#### Test 9: Does price differ by star category? (One-way ANOVA and Tukey test)

* H0: mean log(price) is the same for all star categories.
* If ANOVA is significant, Tukey's HSD shows which pairs differ.

```{r test-anova}
anova_model <- aov(log(price) ~ star_cat, data = price_data)
summary(anova_model)

TukeyHSD(anova_model)

price_data |>
  group_by(star_cat) |>
  summarise(hotels = n(), typical_price = round(exp(mean(log(price)))))
```

```{r test-anova-check}
# Check the ANOVA assumption: are the residuals roughly normal?
qqnorm(residuals(anova_model), main = "Normal Q-Q plot of ANOVA residuals")
qqline(residuals(anova_model), col = "red")

# Non-parametric double-check
kruskal.test(log(price) ~ star_cat, data = price_data)
```

#### Test 10: Are ratings linearly related to price? (Pearson correlation)

* H0: the Pearson correlation between log(price) and the variable is 0.
* We use log(price) because Pearson correlation is sensitive to extreme values.

```{r test-pearson}
pearson_vars <- c("rating_overall", "rating_room", "rating_value", "star_num")

pearson_table <- map_dfr(pearson_vars, function(v) {
  ok <- complete.cases(price_data$price, price_data[[v]])
  t  <- cor.test(log(price_data$price[ok]), price_data[[v]][ok], method = "pearson")
  tibble(variable = v, n = sum(ok), r = round(unname(t$estimate), 3),
         ci_low = round(t$conf.int[1], 3), ci_high = round(t$conf.int[2], 3),
         p_value = signif(t$p.value, 3))
})

pearson_table
```

**Interpretation of Tests 7 to 10:**

* Test 7 (t-test, ratings): t = 12.1 and p < 0.001, so we reject H0. Competitors average 3.93 and Ramada 3.01; the 95% confidence interval says the true gap is between 0.78 and 1.08 stars. This confirms the Wilcoxon result from Test 4: Ramada is clearly rated lower.

* Test 8 (t-test, price): p < 0.001. The typical Hamilton hotel charges about $161 compared with about $204 elsewhere, so Hamilton is roughly 21% cheaper (the confidence interval on the log scale means somewhere between 12% and 29% cheaper). Ramada's $183.50 is above the typical Hamilton price but below the typical price in other areas.

* Test 9 (ANOVA): F = 19.9 and p < 0.001, so price differs across star categories. Tukey's test shows where: 4-5 star hotels (typical price $261) cost about 50% more than 3-3.5 star hotels ($174) and about 94% more than 1-2.5 star hotels ($134), and both differences are significant. The 1-2.5, 3-3.5 and Unrated groups are not significantly different from each other (p between 0.10 and 0.96). So the big price jump only comes at the 4-star level. Ramada (3.5 stars) belongs to the 3-3.5 group, and $183.50 is about 5% above that group's typical price. The residuals are roughly normal in the middle of the plot but not perfectly so in the tails, which is acceptable for a sample of 435 hotels, and the non-parametric Kruskal-Wallis check gives the same conclusion.

* Test 10 (Pearson): log(price) has a moderate positive linear correlation with star class (r = 0.54), the room rating (0.39), the overall rating (0.35) and the value rating (0.24). All p-values are far below 0.05. These correlations are stronger than the raw-price correlations in the heatmap (for example 0.35 instead of 0.22 for the overall rating) because the log transformation stops a few extreme prices from dominating. Star class and room quality are what the market pays for most; the value rating is the weakest link to price.

## Dimensionality reduction: Principal Component Analysis (PCA) of the ratings

We have six rating scores for each hotel (cleanliness, location, room, service, sleep, value), and the correlation heatmap showed they overlap a lot. PCA combines correlated variables into a few new variables called principal components (PCs) that capture most of the information. PC1 captures the largest possible share of the variation, PC2 the next largest, and so on. We standardize the ratings first (`scale. = TRUE`) so that each counts equally. Ramada is included because it is in `hotel_pred`.

```{r pca-run}
# Hotels with all six ratings
pca_input <- hotel_all |>
  select(id, name, locality, rating_cleanliness, rating_location,
         rating_room, rating_service, rating_sleep, rating_value) |>
  drop_na()

cat("Hotels used for PCA:", nrow(pca_input), "\n")

pca_result <- prcomp(pca_input |> select(starts_with("rating_")), scale. = TRUE)
summary(pca_result)
```

```{r pca-scree}
# How much information does each component keep?
var_explained <- tibble(component = paste0("PC", 1:6),
                        pct = pca_result$sdev^2 / sum(pca_result$sdev^2) * 100)

ggplot(var_explained, aes(x = component, y = pct)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(round(pct, 1), "%")), vjust = -0.3) +
  labs(title = "Scree plot: share of variation explained by each component",
       x = NULL, y = "% of variation")
```

```{r pca-loadings}
# Loadings show how much each rating contributes to a component.
# The sign of a PC is arbitrary, so we reverse it so that a HIGHER score means BETTER ratings.
loadings <- data.frame(
  rating = rownames(pca_result$rotation),
  round(-pca_result$rotation[, 1:2], 2)
)

loadings
```

```{r pca-scores}
pca_scores <- as_tibble(-pca_result$x[, 1:2]) |>
  bind_cols(pca_input |> select(id, name, locality)) |>
  mutate(group = case_when(id == ramada_id ~ "Ramada",
                           locality == "Hamilton" ~ "Hamilton competitors",
                           TRUE ~ "Other areas"))

ggplot(pca_scores, aes(x = PC1, y = PC2, colour = group)) +
  geom_point(data = filter(pca_scores, group != "Ramada"), alpha = 0.6) +
  geom_point(data = filter(pca_scores, group == "Ramada"), size = 5) +
  scale_colour_manual(values = c("Ramada" = "firebrick", "Hamilton competitors" = "steelblue",
                                 "Other areas" = "grey70")) +
  labs(title = "Hotels on the first two principal components",
       subtitle = "Right = better overall ratings; Up = location is strong compared with other ratings",
       x = "PC1 (overall quality score)", y = "PC2 (location vs other aspects)", colour = NULL)
```

```{r pca-ramada}
ram <- pca_scores |> filter(group == "Ramada")
ham <- pca_scores |> filter(group == "Hamilton competitors")

cat("Ramada PC1:", round(ram$PC1, 2), "| Hamilton competitors average PC1:", round(mean(ham$PC1), 2), "\n")
cat("Share of Hamilton competitors with a BETTER PC1 score than Ramada:",
    round(mean(ham$PC1 > ram$PC1) * 100), "%\n")
cat("Ramada PC2:", round(ram$PC2, 2), "| Hamilton competitors average PC2:", round(mean(ham$PC2), 2), "\n")

# Does the PC1 quality score relate to price?
pca_price <- pca_scores |>
  left_join(hotel_all |> select(id, price), by = "id") |>
  filter(!is.na(price))

cor.test(pca_price$PC1, log(pca_price$price))
```

**Interpretation:**

* Two components are enough. PC1 keeps 82.7% of the information in the six ratings and PC2 another 6.8%, so together they keep 89.4%. The scree plot drops sharply after PC1.

* What the components mean. PC1 gives almost equal weight (0.36 to 0.42) to all six ratings, so it is an overall quality score. PC2 is dominated by location (0.91) with small opposite weights on the other ratings, so it measures how strong the location is compared with the other aspects.

* Ramada's position. Ramada's quality score (PC1) is -2.79 compared with an average of -1.41 for Hamilton competitors, and 75% of Hamilton competitors score better. On PC2 Ramada scores 0.87 compared with about -0.08 for competitors, which means its location rating is unusually strong relative to its other ratings. On the plot Ramada sits to the left of most hotels (lower overall quality) and above the cloud (strong location). This matches the earlier findings.

* Link to price. The quality score has a significant correlation with log price (r = 0.34, p < 0.001): hotels with better overall ratings charge more.

* Why PCA matters? The six ratings overlap so much that one or two components can replace them in a model and avoid the multicollinearity problem. We kept the original ratings in Phase 4 because they are easier to explain to the manager.

## Cluster analysis: who are Ramada's true peers?

Comparing Ramada with all Hamilton hotels mixes backpackers, motels and large hotels. K-means clustering groups hotels that are similar on price, overall rating, value rating and number of rooms, so we can see which group Ramada really belongs to. Variables are standardized first so that price (in dollars) does not dominate. We use Ramada's current price of $183.50 (from the brief) because `hotel_pred` has no price.

```{r cluster-prepare}
cluster_data <- hotel_all |>
  filter(locality == "Hamilton") |>
  mutate(price = if_else(id == ramada_id, 183.5, price)) |>
  select(id, name, category, price, rating_overall, rating_value, room) |>
  drop_na()

cat("Hamilton hotels with complete data:", nrow(cluster_data), "\n")

cluster_scaled <- scale(cluster_data |> select(price, rating_overall, rating_value, room))
```

**How many clusters?** The "elbow" plot shows how tightly packed the clusters are for 1 to 8 clusters. We look for the point where adding clusters stops helping much.

```{r cluster-elbow}
set.seed(123)
wss <- sapply(1:8, function(k) kmeans(cluster_scaled, centers = k, nstart = 25)$tot.withinss)

tibble(k = 1:8, wss = wss) |>
  ggplot(aes(x = k, y = wss)) +
  geom_line() + geom_point(size = 2) +
  labs(title = "Elbow plot", x = "Number of clusters (k)", y = "Total within-cluster variation")
```

```{r cluster-run}
set.seed(123)
km <- kmeans(cluster_scaled, centers = 3, nstart = 25)

cluster_data$cluster_raw <- km$cluster

# Number the groups by average rating (Group 1 = lowest rated) so the labels are easy to read
cluster_order <- cluster_data |>
  group_by(cluster_raw) |>
  summarise(avg_rating = mean(rating_overall)) |>
  arrange(avg_rating) |>
  mutate(group_no = row_number())

cluster_data <- cluster_data |>
  left_join(cluster_order |> select(cluster_raw, group_no), by = "cluster_raw") |>
  mutate(cluster = factor(paste("Group", group_no), levels = paste("Group", 1:3)))

# Profile of each group
cluster_data |>
  group_by(cluster) |>
  summarise(hotels = n(), avg_price = round(mean(price)), avg_rating = round(mean(rating_overall), 2),
            avg_value_rating = round(mean(rating_value), 2), avg_rooms = round(mean(room)))
```

```{r cluster-ramada}
ramada_group <- cluster_data |> filter(id == ramada_id) |> pull(cluster)
cat("Ramada belongs to:", as.character(ramada_group), "\n\n")

# Ramada's true peers
cluster_data |>
  filter(cluster == ramada_group) |>
  arrange(desc(price)) |>
  select(name, category, price, rating_overall, rating_value, room)
```

```{r cluster-plot}
ggplot(cluster_data, aes(x = rating_overall, y = price, colour = cluster, size = room)) +
  geom_point(alpha = 0.7) +
  geom_point(data = filter(cluster_data, id == ramada_id), colour = "black", shape = 1, stroke = 1.5) +
  geom_text(data = filter(cluster_data, id == ramada_id), label = "Ramada", colour = "black",
            vjust = -1.5, size = 4, show.legend = FALSE) +
  labs(title = "Hamilton hotel groups (circle = Ramada)",
       x = "Overall rating", y = "Price per night (NZD)", colour = "Group", size = "Rooms")
```

**Interpretation:**

* The elbow plot declines smoothly without a sharp bend, so the choice is partly a judgement. Three groups give a clear and useful story, but the result should be seen as a description, not a precise rule. Only 50 Hamilton hotels have complete data.

* Group 1 (17 hotels): average price $136, rating 3.12, value rating 3.24, about 30 rooms. Mostly motels and lodges with average guest satisfaction.

* Group 2 (3 hotels): large properties (about 163 rooms), price $236, rating 3.73.

* Group 3 (30 hotels): price $206, rating 4.39, value rating 4.47, small (about 21 rooms). These are the best-rated hotels.

* Ramada belongs to Group 1, so its real peers are lower-rated motels and lodges, not the top-rated hotels. Within this group, Ramada's price of $183.50 is the second highest (only Sails Motor Inn at $293 is higher) and about $48 (35%) above the group average of $136, while its rating (3.0) and value rating (3.3) are about the same as the group's (3.12 and 3.24).

* By contrast, guests of the high-rated Group 3 pay only about $22 (11%) more than Ramada's price but rate their hotels 4.39.

* Ramada charges almost the price of the top group for the satisfaction level of its peer group. This explains its weak value-for-money score and is another reason not to raise the price.

# 5. Strengths and Weaknesses of Ramada from Reviews

In this phase we mine the written reviews using n-gramS, trying n = 1, 2, 3, 4, 5 and 6 one after the other. Star ratings tell us how guests felt; the words tell us why.

An n-gram is a window of n consecutive words that slides through a review one word at a time. For the sentence "the bathroom door would not close properly":

Small n (1 or 2): the same words and phrases appear in many reviews, so counts are reliable, but there is little context.

* Large n (5 or 6): the phrase carries a lot of context, including words like "not" that change the meaning, but it is almost never written the same way twice.

We go gradually from n = 1 to n = 6 and, at each step, check what we gain and what we lose.

**Steps in this phase**

1. Prepare the text and split it into words
2. Stemming and lemmatization
3. Build the n-grams for n = 1 to 6 and measure how often they repeat
4. Keep the meaningful n-grams, then draw bar diagrams for n = 1 to 6
5. Draw word clouds for n = 1 to 6
6. Topic analysis (LDA), with phrases for each topic at n = 2 to 6
7. Sentiment analysis for n = 1 to 6
8. Aspect analysis, logistic regression and a summary of strengths and weaknesses

## 5.1 Preparing the review text

Computers cannot read sentences, so we first split each review into single words (tokenisation). `unnest_tokens()` also converts the words to lower case and removes punctuation. We keep only real words (letters and apostrophes, so that "wasn't" stays one word), because the word "not" matters for sentiment.

We do not remove stop words (such as "the", "and", "was") at this stage. Inside an n-gram these small words keep the phrase readable. Later, we keep only the n-grams that also contain meaningful words.

```{r text-prep}
# Words that carry little meaning here (used later for topic analysis only)
custom_stop <- c("hotel", "ramada", "wyndham", "hamilton", "stay", "stayed", "staying",
                 "night", "nights", "nz", "city", "centre", "center", "sudima")

# One row per word, in the original order of each review
tokens_words <- ramada_reviews |>
  select(review_id, rating_overall, sentiment_group, full_text) |>
  unnest_tokens(word, full_text) |>
  filter(str_detect(word, "^[a-z']+$"))        # keep real words only

cat("Ramada reviews:", n_distinct(tokens_words$review_id), "\n")
cat("Total words:", nrow(tokens_words), "\n")
cat("Different words:", n_distinct(tokens_words$word), "\n")
cat("Average words per review:", round(nrow(tokens_words) / n_distinct(tokens_words$review_id)), "\n")

# The first 15 words of the first review
head(tokens_words$word, 15)
```

## 5.2 Stemming and lemmatization

Guests write "room", "rooms", "stay", "stayed", "staying" and so on. For a computer these are different words, so identical phrases written with different word forms would not match. We reduce each word to a common base form before building the n-grams.

* **Stemming** chops the end off a word using simple rules. It is fast, but the result may not be a real word.

* **Lemmatization** uses a dictionary to return the proper base word ("rooms" becomes "room", "stayed" becomes "stay", "was" becomes "be"). It sometimes gives odd results for words that look like another word (for example "bit" can become "bite").

```{r stem-lemma-demo}
demo_words <- c("rooms", "cleaned", "cleaning", "stayed", "staying", "was", "were","friendly", "noisy", "better", "walking")

tibble(original = demo_words,
       stem     = wordStem(demo_words),
       lemma    = lemmatize_words(demo_words))
```

```{r stem-lemma-apply}
# Add the stem and the lemma of every word
tokens_words <- tokens_words |>
  mutate(stem  = wordStem(word),
         lemma = lemmatize_words(word))

# How much smaller does the vocabulary become?
vocab_size <- tibble(
  version = c("Original words", "After stemming", "After lemmatization"),
  different_words = c(n_distinct(tokens_words$word),
                      n_distinct(tokens_words$stem),
                      n_distinct(tokens_words$lemma))
) |>
  mutate(reduction_pct = round((1 - different_words / first(different_words)) * 100, 1))

vocab_size
```

```{r vocab-plot}
ggplot(vocab_size, aes(x = factor(version, levels = version), y = different_words)) +
  geom_col(fill = "firebrick") +
  geom_text(aes(label = paste0(different_words, " (-", reduction_pct, "%)")), vjust = -0.3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(title = "Stemming and lemmatization shrink the vocabulary",
       x = NULL, y = "Number of different words")
```

**Interpretation:**

* Ramada's 349 reviews contain about 32,600 words (about 93 per review), but only 3,338 different word forms.

* Stemming reduces the vocabulary by about 23% (to 2,581) and lemmatization by about 22% (to 2,608). Both group different forms of the same word, so identical ideas are counted together.

* Stems are hard to read ("stayed" becomes "stai", "was" becomes "wa", "friendly" becomes "friendli"). Lemmas are real words ("stayed" becomes "stay", and "was" and "were" both become "be", "better" becomes "good"). For readable charts we therefore use the lemmatized text.

* Lemmatization is not perfect: it can give odd results such as "bit" becoming "bite" or "bubbles" becoming "bubble". Because "was", "is" and "were" all become "be", you will see phrases such as "the room be clean and tidy". Read "be" as "was/is/were".

## 5.3 Building the n-grams for n = 1 to 6

We rebuild each review as one string of words in three versions (original, stemmed and lemmatized) and then slide a window of n words through it with `unnest_tokens(token = "ngrams", n = n)`. Windows never cross from one review into the next. A review shorter than n words produces no n-gram.

For every n and every version we count:

* Different n-grams: how many distinct word sequences there are, and 

* repeated n-grams: how many appear in two or more different reviews. We count each n-gram once per review, so a guest who repeats a phrase in the same review does not inflate the count.

```{r build-ngrams}
ngram_names <- c("single words (1-grams)", "2-grams", "3-grams", "4-grams", "5-grams", "6-grams")

# One text per review for each version of the words
review_texts <- tokens_words |>
  group_by(review_id, rating_overall, sentiment_group) |>
  summarise(text_original = paste(word,  collapse = " "),
            text_stem     = paste(stem,  collapse = " "),
            text_lemma    = paste(lemma, collapse = " "),
            .groups = "drop")

# Count different and repeated n-grams for one value of n and one version of the words
count_repeats <- function(n, version) {
  text_col <- c(original = "text_original", stem = "text_stem", lemma = "text_lemma")[[version]]

  grams <- review_texts |>
    select(review_id, text = all_of(text_col)) |>
    unnest_tokens(ngram, text, token = "ngrams", n = n) |>
    filter(!is.na(ngram))

  per_gram <- grams |>
    group_by(ngram) |>
    summarise(n_reviews = n_distinct(review_id), .groups = "drop")

  tibble(n = n, version = version, total = nrow(grams),
         different = nrow(per_gram), repeated = sum(per_gram$n_reviews >= 2))
}

# Run it for n = 1 to 6 and the three versions
ngram_summary <- expand_grid(n = 1:6, version = c("original", "stem", "lemma")) |>
  pmap_dfr(count_repeats)

# One tidy table: counts for the lemmatized version, and the % of n-grams that repeat for each version
repeat_pct <- ngram_summary |>
  mutate(pct_repeated = round(repeated / different * 100, 2)) |>
  select(n, version, pct_repeated) |>
  pivot_wider(names_from = version, values_from = pct_repeated, names_prefix = "pct_repeated_")

repeat_table <- ngram_summary |>
  filter(version == "lemma") |>
  select(n, total_ngrams = total, different_ngrams = different, repeated_lemma = repeated) |>
  left_join(repeat_pct, by = "n") |>
  select(n, total_ngrams, different_ngrams, repeated_lemma,
         pct_repeated_original, pct_repeated_stem, pct_repeated_lemma) |>
  mutate(pct_repeated = pct_repeated_lemma)        # used by the plot below

repeat_table |> select(-pct_repeated)
```

```{r repeat-plot}
ggplot(repeat_table, aes(x = n, y = pct_repeated)) +
  geom_line(colour = "firebrick", linewidth = 1) +
  geom_point(size = 3, colour = "firebrick") +
  geom_text(aes(label = paste0(pct_repeated, "%")), vjust = -1) +
  scale_x_continuous(breaks = 1:6) +
  scale_y_log10(expand = expansion(mult = c(0.1, 0.4))) +
  labs(title = "The longer the n-gram, the fewer repeats",
       subtitle = "% of different lemmatized n-grams that appear in 2 or more reviews (log scale)",
       x = "n (number of words in the n-gram)", y = "% of n-grams that repeat (log scale)")
```

**Interpretation:**

* Repetition falls quickly as n grows.Of the different lemmatized n-grams, 53.9% of single words appear in two or more reviews, but only 22.1% of 2-grams, 7.8% of 3-grams, 2.2% of 4-grams, 0.6% of 5-grams and just 0.18% of 6-grams. Each extra word cuts the repeat rate by a factor of roughly 2.5 to 3.7. At n = 6, 54 of 30,782 different 6-grams repeat.

* The number of different n-grams explodes: from about 2,600 single words to about 30,800 different 6-grams from roughly the same amount of text (about 31,000 to 32,000 windows for every n). By n = 6 almost every window is unique.

* Stemming and lemmatization always help, and help more as n grows. Compared with the original words, lemmatization raises the repeat rate by about 8% for single words, 11% for 2-grams, 17% for 3-grams, 21% for 4-grams, 25% for 5-grams and 29% for 6-grams (for example 43 to 54 repeated 6-grams). Stemming gives similar gains. Long phrases are the most sensitive to word forms ("was friendly" and "were friendly" only match after lemmatization).

* The trade-off in one sentence: small n gives reliable counts but little context; large n gives rich context but only a few repeated phrases. That is why we look at all six values and combine them with topic and sentiment analysis.

## 5.4 Keeping the meaningful n-grams

Many n-grams contain only small words ("and the", "was in a"). These are called stop words. Stop words are useful inside a phrase because they make it readable, but an n-gram made mostly of stop words says nothing. We therefore use one simple rule at every n:

> Keep an n-gram only if it contains **at least two meaningful (non-stop) words**. For single words (n = 1), the word itself must be meaningful.

So "car park" and "friendly and helpful" are kept, while "and the" and "in the" are dropped. We use the lemmatized text, count each n-gram once per review, and also record the average star rating of the reviews that contain it (to see whether the phrase belongs to happy or unhappy guests).

```{r meaningful-ngrams}
# Stop words: the standard list plus our own hotel-specific words
stop_set <- unique(c(stop_words$word, custom_stop))

gram_tables <- map(1:6, function(k) {
  review_texts |>
    select(review_id, rating_overall, text_lemma) |>
    unnest_tokens(ngram, text_lemma, token = "ngrams", n = k) |>
    filter(!is.na(ngram)) |>
    distinct(review_id, ngram, rating_overall) |>         # count once per review
    group_by(ngram) |>
    summarise(n_reviews = n(), avg_rating = round(mean(rating_overall), 2), .groups = "drop") |>
    # how many words in the n-gram are meaningful?
    mutate(meaningful_words = rowSums(matrix(!(str_split_fixed(ngram, " ", k) %in% stop_set), ncol = k))) |>
    filter(meaningful_words >= min(k, 2)) |>
    arrange(desc(n_reviews), avg_rating)
})

# Summary for n = 1 to 6
meaningful_summary <- map_dfr(1:6, function(k) {
  tb <- gram_tables[[k]]
  tibble(n = k,
         different_meaningful = nrow(tb),
         repeated_in_2_or_more_reviews = sum(tb$n_reviews >= 2),
         most_common = tb$ngram[1],
         in_how_many_reviews = tb$n_reviews[1])
})

meaningful_summary
```

```{r top-ngram-tables}
# The top 10 repeated meaningful n-grams for each n, side by side
map_dfr(1:6, function(k) {
  gram_tables[[k]] |>
    filter(n_reviews >= 2) |>
    slice_head(n = 10) |>
    transmute(n = k, rank = row_number(), ngram, reviews = n_reviews, avg_rating)
}) |>
  print(n = Inf)
```

## 5.5 Bar diagrams for n = 1 to 6

A bar diagram shows the most repeated n-grams. Each bar is the number of different reviews that contain the n-gram. The colour shows the mood of the guests who used it: 

green = happy guests (average rating 3.5 or more), 
red   = unhappy guests (average rating 2.5 or less), 
grey  = nutral. 

We draw one diagram for each n, from single words up to 6-grams, so you can see how the picture changes as the phrases get longer.

```{r bar-diagrams, fig.height=5.5, fig.width=9}
mood_colours <- c("Happy guests (avg rating 3.5+)"        = "darkgreen",
                  "Unhappy guests (avg rating 2.5 or less)" = "firebrick",
                  "nutral"                                  = "grey60")

add_mood <- function(tb) {
  tb |> mutate(mood = case_when(avg_rating >= 3.5 ~ "Happy guests (avg rating 3.5+)",
                                avg_rating <= 2.5 ~ "Unhappy guests (avg rating 2.5 or less)",
                                TRUE              ~ "nutral"))
}

for (k in 1:6) {
  p <- gram_tables[[k]] |>
    filter(n_reviews >= 2) |>
    add_mood() |>
    slice_head(n = 15) |>
    ggplot(aes(x = reorder(ngram, n_reviews), y = n_reviews, fill = mood)) +
    geom_col() +
    coord_flip() +
    scale_fill_manual(values = mood_colours) +
    labs(title = paste0("Top repeated ", ngram_names[k], " (lemmatized, meaningful)"),
         x = NULL, y = "Number of reviews containing it", fill = NULL) +
    theme(legend.position = "bottom")
  print(p)
}
```

**Interpretation:**

* n = 1 (single words): the most widespread words are *park* (138 reviews), *staff* (130), *clean* (128), *location* (127), *bed* (105), *bathroom* (93), *restaurant* (81), *book* (75) and *shower* (74). The average rating of the reviews tells us more than the count: *bathroom* (2.5), *shower* (2.6), *bed* (2.7) and *book* (2.8) go with unhappy guests, while *friendly* (3.7) and *restaurant* (3.5) go with happier ones. Words like *staff*, *park* and *location* are used by both happy and unhappy guests, so a single word cannot tell us which.

* **n = 2:** pairs add the missing context: **car park** (36 reviews, average rating 3.1), **central location** (22, 3.8), **main street** (16, 3.3), **front desk** (14, 2.9), **walk distance** (14, 4.1), **air condition** (13, 2.5), **customer service** (11, 3.1), **friendly staff** (11, 4.0) and **rabbit warren** (9, 2.4). Now the phrases separate: central location, walking distance and friendly staff are green; air conditioning and the rabbit-warren layout are red.

* **n = 3:** *friendly and helpful* (17 reviews), *staff be friendly* (15), *clean and tidy* (13) and *restaurant and bar* (10) are green, while *the front desk* (2.4), *the car park* (2.7) and *the main street* (2.8) are the unhappy ones.

* **n = 4:** *staff be friendly and* (11 reviews) and *be friendly and helpful* (10) are the most repeated, followed by unhappy phrases such as *on the main street* (2.6), ***from the car park*** (2.3) and ***step into the shower*** (2.0).

* **n = 5 and n = 6:** only 59 and 23 meaningful phrases repeat, but they read like real sentences: *staff be friendly and helpful* (9 reviews, 4.1), *room be clean and tidy*, *have to pay for park* (2.7), *get from the car park to* (2.7), *place be like a rabbit warren* (2.0), *room be on the main street* (2.5) and *step to get into the shower* (2.5).

* **Following one theme as n grows:** the same complaint becomes more specific. *car park* (36 reviews) becomes *from the car park* (6) and then *get from the car park to* (3): guests complain about **getting from the car park to the hotel**. *shower* (74) becomes *step into the shower* (4) and then *step to get into the shower* (2): the problem is the **high step into the shower**. The positive theme follows the same path: *friendly staff* becomes *staff be friendly and helpful*.

* Remember that "be" is the lemma of "was/is/were", so "staff be friendly" means "staff were friendly".

## 5.6 Word clouds for n = 1 to 6

A word cloud shows the repeated n-grams at a glance: bigger text means the n-gram appears in more reviews. The colours are the same as in the bar diagrams (green = happy guests, red = unhappy guests, grey = nutral). Longer n-grams need smaller text to fit, so we show fewer phrases as n grows.

```{r wordclouds-ngrams, fig.height=6.5, fig.width=10}
cloud_max   <- c(100, 80, 60, 50, 40, 30)
cloud_scale <- list(c(4, 0.7), c(3, 0.6), c(2.2, 0.5), c(1.6, 0.4), c(1.3, 0.4), c(1.1, 0.4))

for (k in 1:6) {
  cd <- gram_tables[[k]] |>
    filter(n_reviews >= 2) |>
    add_mood() |>
    arrange(desc(n_reviews), avg_rating) |>
    slice_head(n = cloud_max[k]) |>
    mutate(colour = case_when(str_detect(mood, "^Happy") ~ "darkgreen",
                              str_detect(mood, "^Unhappy") ~ "firebrick",
                              TRUE ~ "grey40"))

  set.seed(123)
  par(mar = c(0, 0, 2, 0))
  wordcloud(words = cd$ngram, freq = cd$n_reviews, min.freq = 2, max.words = cloud_max[k],
            scale = cloud_scale[[k]], rot.per = 0, random.order = FALSE,
            ordered.colors = TRUE, colors = cd$colour)
  title(paste0("Word cloud of repeated ", ngram_names[k],
               " (green = happy guests, red = unhappy guests, grey = nutral)"), cex.main = 0.9)
}
```

**Interpretation:**

* n = 1: the biggest words (location, bed, clean, park, staff, bathroom, restaurant, shower, book) are mostly **grey**, because happy and unhappy guests both use them. Colour appears in the smaller words: **red** for door, smell, noisy, window, pay, dirty, disappoint, terrible, **green** for friendly, helpful, comfortable, central, restaurant, spacious, quiet. Single words show what guests talk about, but not whether the comment is good or bad.

* n = 2 to 4: the colours become much clearer, because a phrase carries context. In the 4-gram cloud the **green** phrases are staff be friendly and helpful, be clean and tidy, within easy walk distance, free park on site and excellent value for money. The **red** phrases are from the car park, step into the shower, bathroom be ridiculously small, to pay for wifi, don't waste your money, a lot of money, pack up and leave, hole in the wall and the main entrance door.

* n = 5 and n = 6: the clouds are smaller, because fewer phrases repeat, but the phrases are almost complete sentences. The same messages appear: friendly staff and clean rooms (green); getting from the car park, the shower step and the rabbit-warren layout (red).

* Overall: word clouds are most useful for n = 2 to 4: the phrases are frequent enough to be reliable and long enough to show whether guests are happy or unhappy.

# 6. Topic analysis (LDA)

Reading hundreds of reviews to find the main themes is slow. Topic analysis finds them automatically. We use Latent Dirichlet Allocation (LDA), which assumes that:

* each review is a mixture of topics (for example 70% "room problems" and 30% "location"), and

* each topic is a group of words that tend to appear together.

LDA does not now the meaning of a topic. It only returns a group of words, and we interpret them. LDA works on single words (n = 1), because longer n-grams are too rare to build topics from. We use the lemmatized words with stop words removed and drop words used in fewer than 5 reviews to remove noise. Afterwards, we use the repeated n-grams for n = 2 to 6 to show what each topic sounds like as phrases.

```{r lda-prepare}
# Meaningful words only
content_words <- tokens_words |>
  filter(!word %in% stop_words$word, !lemma %in% stop_words$word,
         !word %in% custom_stop, !lemma %in% custom_stop, nchar(lemma) > 2)

# Keep words used in at least 5 reviews
common_terms <- content_words |>
  distinct(review_id, lemma) |>
  count(lemma) |>
  filter(n >= 5) |>
  pull(lemma)

# Document-term matrix: one row per review, one column per word
dtm <- content_words |>
  filter(lemma %in% common_terms) |>
  count(review_id, lemma) |>
  cast_dtm(document = review_id, term = lemma, value = n)

cat("Reviews:", nrow(dtm), "| Words:", ncol(dtm), "\n")
```

We try 2 to 6 topics. Perplexity measures how well the model predicts the words (lower is better), but it keeps falling slowly as topics are added, so we also look for topics that are easy to interpret. We choose 4.

```{r lda-choose-k}
perplexity_table <- map_dfr(2:6, function(k) {
  m <- LDA(dtm, k = k, method = "Gibbs", control = list(seed = 123, burnin = 200, iter = 1000))
  tibble(topics = k, perplexity = round(perplexity(m, dtm), 1))
})

perplexity_table
```

```{r lda-fit}
lda_model <- LDA(dtm, k = 4, method = "Gibbs", control = list(seed = 123, burnin = 200, iter = 1000))

# Top 10 words in each topic (beta = how strongly a word belongs to a topic)
topic_words <- tidy(lda_model, matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 10, with_ties = FALSE) |>
  ungroup()

topic_labels <- topic_words |>
  group_by(topic) |>
  summarise(label = paste0("Topic ", first(topic), ": ", paste(term[1:3], collapse = ", ")))

topic_words |>
  group_by(topic) |>
  summarise(top_words = paste(term, collapse = ", "))
```

```{r lda-plot-words, fig.height=7}
topic_words |>
  left_join(topic_labels, by = "topic") |>
  ggplot(aes(x = reorder_within(term, beta, label), y = beta, fill = label)) +
  geom_col(show.legend = FALSE) +
  scale_x_reordered() +
  coord_flip() +
  facet_wrap(~ label, scales = "free_y") +
  labs(title = "The four topics in Ramada reviews: top 10 words", x = NULL, y = "Word weight in the topic")
```

We assign each review to the topic with the highest share (gamma) and then look at the ratings.

```{r lda-topic-ratings}
dominant_topic <- tidy(lda_model, matrix = "gamma") |>
  mutate(review_id = as.numeric(document)) |>
  group_by(review_id) |>
  slice_max(gamma, n = 1, with_ties = FALSE) |>
  ungroup() |>
  left_join(ramada_reviews |> select(review_id, rating_overall), by = "review_id")

topic_summary <- dominant_topic |>
  group_by(topic) |>
  summarise(reviews = n(),
            share_of_reviews_pct = round(n() / nrow(dominant_topic) * 100),
            avg_rating = round(mean(rating_overall), 2),
            pct_negative_reviews = round(mean(rating_overall <= 2) * 100)) |>
  left_join(topic_labels, by = "topic")

topic_summary |> select(label, reviews, share_of_reviews_pct, avg_rating, pct_negative_reviews)
```

```{r lda-plot-ratings}
ggplot(topic_summary, aes(x = reorder(label, avg_rating), y = avg_rating, fill = avg_rating)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(avg_rating, " (", pct_negative_reviews, "% negative)")), hjust = -0.05, size = 3.5) +
  coord_flip(ylim = c(0, 5.5)) +
  scale_fill_gradient(low = "firebrick", high = "darkgreen") +
  labs(title = "Average star rating of the reviews in each topic", x = NULL, y = "Average overall rating")
```

For each topic and each n from 2 to 6, we pick the most repeated n-gram that contains one of the topic's top-10 words. Reading across a row shows how the same theme grows from a short phrase into a full expression.

```{r lda-phrases}
topic_phrases <- map_dfr(1:4, function(t) {
  terms   <- topic_words |> filter(topic == t) |> pull(term)
  pattern <- paste0("\\b(", paste(terms, collapse = "|"), ")\\b")

  map_dfr(2:6, function(k) {
    gram_tables[[k]] |>
      filter(n_reviews >= 2, str_detect(ngram, pattern)) |>
      slice_head(n = 1) |>
      mutate(n = k, topic = t)
  })
}) |>
  left_join(topic_labels, by = "topic")

topic_phrases |>
  select(label, n, ngram) |>
  pivot_wider(names_from = n, names_prefix = "n = ", values_from = ngram)
```

**Interpretation:**

* Choice of topics: perplexity falls from about 265 (2 topics) to 254 (4 topics) and then flattens at about 250 (5 and 6 topics). Four topics are easy to interpret, so we use them.

* Topic made of clean, staff, location, restaurant, comfortable, friendly, helpful is the happy-guest topic: about 30% of reviews, average rating 3.96, only 12% negative. Its phrases grow from central location to friendly and helpful, staff be friendly and helpful and the room be clean and tidy.

* Topic made of bed, bathroom, shower, door, smell, window, sleep, people, step, noise is the problem topic: about 29% of reviews, average rating just 1.99, and 71% negative. Its phrases end with step to get into the shower: the high shower step is a clear complaint. (The phrase comfy bed is also picked for this topic simply because it contains the word "bed"; the topic mixes some praise with many complaints, but most of its reviews are negative.)

* Topic made of park, car, reception, central, street, pay, free, front is about arrival, parking and reception: about 24% of reviews, average rating 3.07, 28% negative. Its phrases are car park, the front desk, on the main street and get from the car park to.

* Topic made of book, check, expect, time, arrive, breakfast, service is about booking, check-in and expectations: about 17% of reviews, average rating 2.97, 36% negative. Its phrases (walk distance, within easy walk distance, check in and check out) show a mix of the good (location) and the practical (check-in), and no phrase repeats at n = 6.

* Reading across a row shows how a theme becomes a full expression as n grows, which is the benefit of longer n-grams. LDA itself needs single words, since longer n-grams are too rare.

* Meaning for the manager: nearly three in ten guests write mainly about room, bathroom and noise problems and rate the hotel about 2 out of 5. Fixing those issues is the most valuable improvement, followed by parking and the check-in experience.

* *Note:* LDA is a statistical model: the topic numbers can change if the model is rerun with a different random seed, but the groups of words stay similar.

